\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-01 paper.qmd
Add true model notation to the table

>>>>

The S-curve model, denoted as \( T \), is simulated through a multi-step process. Initially, a random variable \( T \) is generated from a uniform distribution between \(-\frac{\pi}{2}\) and \(\frac{\pi}{2}\), representing the parameter \( T \) in the S-curve model. Subsequently, three variables \( x_1 \), \( x_2 \), and \( x_3 \) are computed based on \( T \) to construct the S-curve data. \( x_1 \) is calculated as the sine of \( T \), \( x_2 \) as twice the value of another uniformly distributed random variable between 0 and 1, and \( x_3 \) as the cosine of \( T \) minus 1, multiplied by the sign of \( T \). Additionally, four noise dimensions are introduced to the dataset to introduce variability: \( x_4 \), \( x_5 \), \( x_6 \), and \( x_7 \), each generated from uniform distributions within specific intervals. Finally, these variables are aggregated to form the S-curve dataset \( X \).

Let \( T \) be a random variable uniformly distributed in the interval \( [-\frac{\pi}{2}, \frac{\pi}{2}] \), i.e., \( T \sim U\left(-\frac{\pi}{2}, \frac{\pi}{2}\right) \).

Then, we define the following variables:

- \( x_1 = \sin(3 \pi T) \)
- \( x_2 = 2.0 \times U(0, 1) \), where \( U(0, 1) \) represents a uniform distribution between 0 and 1.
- \( x_3 = \text{sign}(T) \times (\cos(3 \pi T) - 1) \)

Additionally, we introduce noise dimensions:

- \( x_4, x_5 \) are uniformly distributed random variables in the interval \( [-0.02, 0.02] \), i.e., \( x_4, x_5 \sim U(-0.02, 0.02) \).
- \( x_6 \) is a uniformly distributed random variable in the interval \( [-0.1, 0.1] \), i.e., \( x_6 \sim U(-0.1, 0.1) \).
- \( x_7 \) is a uniformly distributed random variable in the interval \( [-0.01, 0.01] \), i.e., \( x_7 \sim U(-0.01, 0.01) \).

Finally, we form the data matrix \( X \) by combining \( x_1, x_2, x_3, x_4, x_5, x_6, \) and \( x_7 \) into columns.

### 2024-04-01 paper.tex
Add true model notation to the table

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
$X$ & $p$-dimensional data\\
$Y$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
$y_{max}$ & maximum value of the scaled second embedding component\\
\addlinespace
$r_1$, $r_2$ & range of the first and  second embedding component\\
$h_b$, $w_b$ & height and width of the hexagon\\
$a_r$ & aspect ratio of the 2D layout before scaling\\
$h_r$ & hexagon ratio\\
$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
\addlinespace
$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
$q_x$, $q_y$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
\addlinespace
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

>>>>

\begin{tabular}{>{\raggedright\arraybackslash}p{3cm}>{\raggedright\arraybackslash}p{12cm}}
\toprule
\textbf{Notation} & \textbf{Description}\\
\midrule
$n, p, k$ & number of observations, variables, embedding dimension, respectively\\
$X$ & $p$-dimensional data\\
$Y$ & $k$-dimensional layout\\
$P$ & orthonormal basis, generating a $d-dimensional$ linear projection of $p$-dimensional data\\
$T$ & true  model\\
\addlinespace
$y_{max}$ & maximum value of the scaled second embedding component\\
$r_1$, $r_2$ & range of the first and  second embedding component\\
$h_b$, $w_b$ & height and width of the hexagon\\
$a_r$ & aspect ratio of the 2D layout before scaling\\
$h_r$ & hexagon ratio\\
\addlinespace
$s$ & hexagonal size (radius of the outer circle surrounding the hexagon)\\
$(x_s, \ y_s)$ & starting coordinates of the hexagonal grid\\
$q_x$, $q_y$ & buffer amount along the x and y axes\\
$h_s$, $v_s$ & horizontal and vertical distance between adjacent hexagon bin centroids\\
$d_x$, $d_y$ & horizontal spacing and adjusted vertical spacing for hexagonal coordinates\\
\addlinespace
$b_x$, $b_y$ & number of hexagon bins along the x and y axes\\
$b$, $b'$ & total and non-empty hexagon bins in the grid\\
$n_k$ & number of observations within the $k^{th}$ hexagon\\
$(h^k_{x_i}, h^k_{y_i})$ & hexagonal grid coordinates of the $k^{th}$ hexagon\\
$C^{(j)}$ & $j$-dimensional bin centers\\
\bottomrule
\end{tabular}
\endgroup{}

@@

The S-curve model, denoted as ( T ), is simulated through a multi-step
process. Initially, a random variable ( T ) is generated from a uniform
distribution between (-\frac{\pi}{2}) and (\frac{\pi}{2}), representing
the parameter ( T ) in the S-curve model. Subsequently, three variables
( x\_1 ), ( x\_2 ), and ( x\_3 ) are computed based on ( T ) to
construct the S-curve data. ( x\_1 ) is calculated as the sine of ( T ),
( x\_2 ) as twice the value of another uniformly distributed random
variable between 0 and 1, and ( x\_3 ) as the cosine of ( T ) minus 1,
multiplied by the sign of ( T ). Additionally, four noise dimensions are
introduced to the dataset to introduce variability: ( x\_4 ), ( x\_5 ),
( x\_6 ), and ( x\_7 ), each generated from uniform distributions within
specific intervals. Finally, these variables are aggregated to form the
S-curve dataset ( X ).

Let ( T ) be a random variable uniformly distributed in the interval (
{[}-\frac{\pi}{2}, \frac{\pi}{2}{]} ), i.e., ( T
\sim U\left(-\frac{\pi}{2}, \frac{\pi}{2}\right) ).

Then, we define the following variables:

\begin{itemize}
\tightlist
\item
  ( x\_1 = \sin(3 \pi T) )
\item
  ( x\_2 = 2.0 \times U(0, 1) ), where ( U(0, 1) ) represents a uniform
  distribution between 0 and 1.
\item
  ( x\_3 = \text{sign}(T) \times (\cos(3 \pi T) - 1) )
\end{itemize}

Additionally, we introduce noise dimensions:

\begin{itemize}
\tightlist
\item
  ( x\_4, x\_5 ) are uniformly distributed random variables in the
  interval ( {[}-0.02, 0.02{]} ), i.e., ( x\_4, x\_5 \sim U(-0.02, 0.02)
  ).
\item
  ( x\_6 ) is a uniformly distributed random variable in the interval (
  {[}-0.1, 0.1{]} ), i.e., ( x\_6 \sim U(-0.1, 0.1) ).
\item
  ( x\_7 ) is a uniformly distributed random variable in the interval (
  {[}-0.01, 0.01{]} ), i.e., ( x\_7 \sim U(-0.01, 0.01) ).
\end{itemize}

Finally, we form the data matrix ( X ) by combining ( x\_1, x\_2, x\_3,
x\_4, x\_5, x\_6, ) and ( x\_7 ) into columns.
